### 2019-03-24 01-intro.Rmd
added bookdown files

# Introduction {#intro}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, fig.pos = 'h', fig.align = 'center')
library(tidyverse)
library(eechidna)
library(ggplot2)
library(ggthemes)
library(knitr)
```

Australia has changed in many ways over the last two decades. Rising house prices, country-wide improvements in education, an ageing population, and a decline in religious affiliation, are just a few facets of the country's evolving socio-demographic characteristics. At the same time, political power has moved back and forth between the two major parties. In the 2007 and 2010 federal elections, the Australian Labor Party (Labor) was victorious, whereas the 2001, 2004, 2013 and 2016 elections were won by the Liberal National coalition (Liberal). The two-party preferred vote, a measure of support between these two parties, fluctuated between 47.3% and 53.5% (in favour of the Liberal party) over this period. This study explores how electoral characteristics relate to two-party preference, and whether their effects have changed over time. Electoral socio-demographics are derived from the Census, and vote counts are obtained from federal elections.

Joining these two data sources is problematic as there in an inherent asynchronicity in the two events. A Census is conducted by the Australian Bureau of Statistics (ABS) every five years, whereas a federal election (conducted by the Australian Electoral Commission (AEC)) usually occurs every three years. The first problem addressed is that of obtaining appropriate Census data for the 2004, 2007, 2010 and 2013 elections - election years in which a Census does not occur. The predominant approach in previous studies is to join voting outcomes to the nearest Census, without accounting for any temporal differences (see @DavisStimson98, @Stimson06, @Liao09 and @Stimson09). Furthermore, electoral boundaries change regularly, so spatial discrepancies also arise when matching electoral data. To obtain appropriate Census data for these four elections, electoral socio-demographics are imputed using a spatio-temporal imputation that combines areal interpolation [@Goodchild1993] and linear time-interpolation. The process of collecting and wrangling the raw data is outlined in section \@ref(data), and the imputation process is detailed in section \@ref(imputation). All data and associated documentation relating to this procedure are available in the `eechidna` `R` package [@eechidna], providing a resource for future analysis.

Previous work on modelling Australian federal elections have found that aggregate socio-demographics are relatively good predictors of voting outcomes. @Forrest01 does this using multiple regression of the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998. @Stimson06, @Stimson09 and @Stimson12 use principal component analysis of polling booths in the 2001, 2004 and 2007 elections respectively, also finding that socio-demographic characteristics of polling booths are linked to their two-party preferred vote. On the contrary, @Stimson09 models the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, finding that little of swing vote can be explained by Census data. Instead of analyzing a single election in isolation, this paper employs a consistent model framework across six elections so that temporal changes in the effects of socio-demographics can be observed, where each federal elections is modelled with a cross-sectional data set. The use of a regression framework to examine these socio-political relationships over time is seemingly absent from previous Australian studies. It also appears that no study has attempted any type of statistical analysis of socio-demographics in conjunction with voter behaviour in Australia since 2007, making this paper distinctly different from those previous.

The cross-sectional data set for each election consists of the two-party preferred vote (response variable), and socio-demographic variables (explanatory variables) that characterise each electorate. To obtain these cross-sections, socio-demographic variables are first standardized, and then principal components are used to group variables into "factors". To account for the inherent spatial structure of the data, a spatial error model is fit for each election. These steps are discussed in section \@ref(modelling). In section \@ref(results) inference is conducted on the models to see which effects are significant, how effects change over time and which electorates have abnormal voting behaviour.