### 2019-03-24 index.Rmd

--- 
title: "Spatial modelling of the electoral two-party preferred vote in Australia. A study of federal elections between 2001 and 2016 via the `eechidna` `R` package."
author: "Jeremy Forbes, Di Cook & Rob Hyndman"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
classoption: openany
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: rstudio/bookdown-demo
description: "This is an incomplete draft of a paper to submitted to ANZJS."
---