### 2019-03-27 04-results.Rmd
add country wide trend, diversity, cooks distance,

# Results

## Spatial autoregressive parameter

The spatial autoregressive coefficient $\rho$ is positive and significant in only the 2001 and 2016 elections (Figure \@ref(fig:rhovis)), meaning that in these elections, an electorate's political climate was affected by the attitudes of it's neighbours. Conversely, in the other four elections, the spatial effect weakens to become marginally insignficant. In these years, it appears that the spatial component does not explain anything over and above the electoral socio-demographics, meaning electorates effectively acted more independently.

```{r rhovis, fig.cap="Estimates of the spatial autoregressive parameter for each of the six elections, with a 95\\% confidence interval. Only in 2001 and 2016 is there a significant spatial component", out.height='20%'}
rho_df <- bind_rows(glsmod16$rho_df %>% mutate(year = "2016"), 
  glsmod13$rho_df %>% mutate(year = "2013"),
  glsmod10$rho_df %>% mutate(year = "2010"), 
  glsmod07$rho_df %>% mutate(year = "2007"),
  glsmod04$rho_df %>% mutate(year = "2004"), 
  glsmod01$rho_df %>% mutate(year = "2001")) %>% 
  mutate(upper95 = estimate + 1.96*se, lower95 = estimate - 1.96*se)

rho_df %>% ggplot() + 
  geom_point(aes(x = year, y = estimate), size = 3) +
  geom_linerange(aes(x = year, ymin = lower95, ymax = upper95), size = 1.5) + 
  geom_hline(aes(yintercept = 0), alpha = 0.5, size = 1) +
  scale_color_manual(values = c("grey50", "black")) +
  labs(x = "Election year", y = "Estimate of spatial autoregressive parameter") +
  lims(y = c(-1,1)) +
  guides(col = F) +
  coord_fixed(ratio = 1)
```

## Country-wide trend
Since all socio-demographics have been standardized to have a mean of zero and a variance of one, the intercept in each model can be interpreted as the estimated two-party preferred vote for an electorate with mean characteristics[^3]. The baseline of party preference has varied over the elections, with the biggest swing occuring in the 2007 election where the mean electorate shifted more than five percentage points in favour of the Labor party.

[^3]: Mean of all variables aside from Judaism, Indigenous, Islam and Buddhism, where it assumes the mean of the log value.

```{r plotintercept, fig.cap = "Estimated intercept for each election, which represents the two-party preferred vote for an electorate with mean characteristics.", out.height="20%"}
intercept_df <- bind_rows(data.frame(summary(glsmod16)$tTable)[1, ]  %>% mutate(year = "2016"), 
  data.frame(summary(glsmod13)$tTable)[1, ] %>% mutate(year = "2013"),
  data.frame(summary(glsmod10)$tTable)[1, ] %>% mutate(year = "2010"), 
  data.frame(summary(glsmod07)$tTable)[1, ] %>% mutate(year = "2007"),
  data.frame(summary(glsmod04)$tTable)[1, ] %>% mutate(year = "2004"), 
  data.frame(summary(glsmod01)$tTable)[1, ] %>% mutate(year = "2001")) %>% 
  rename(se = `Std.Error`, estimate = Value, p = `p.value`) %>% 
  mutate(upper95 = estimate + 1.96*se, lower95 = estimate - 1.96*se)

intercept_df %>% ggplot() + 
  geom_point(aes(x = year, y = estimate), size = 2) +
  geom_linerange(aes(x = year, ymin = lower95, ymax = upper95), size = 1.5) + 
  geom_hline(aes(yintercept = 50), alpha = 0.5, size = 1) +
  labs(x = "Election year", y = "Estimate of intercept") +
  guides(col = F) +
  coord_fixed(ratio = 0.2) +
  lims(y = c(45, 55))
```

## Influential socio-demographics
To investigate the socio-demographics that have a strong effect on the two-party preferred vote, partial residual plots are used. These show the direction, size and significance of an estimated effect - the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95% confidence band. If a horizontal line can be drawn through the confidence band, then the effect is insignificant. Plots for each election are faceted to compare the effects over time.

*It is important here to note the ecological fallacy - insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g. individual voters).*

### Income and unemployment
Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs. Despite this, incomes have only significantly affected the electoral two-party preference in 2016 (see Figure \@ref(fig:plotincomes)), with higher income areas associated with support for the Liberal party. In the prior elections, the estimated effect is still positive, albeit insignificant. Unemployment on the other hand, was influential in 2001, 2004 and 2007 elections (Figure \@ref(fig:plotunemploy)), with electorates with higher unemployment favouring Labor. This estimated effect has become weaker over time, with unemployment having virtually no effect in 2016.

```{r myvisreg, include = F}
# Function to produce visreg style conditional plots
my_visreg <- function(my_model, sp_weights, varname, 
  nolabs = FALSE, xlimits = NULL, ylimits = NULL, year = "") {

  # Extract fitted parameters
  rho <- my_model$rho_df$estimate
  sigma <- sqrt(sum(my_model$residuals^2)/(my_model$dims$N-my_model$dims$p))

  # Spatial weights
  w_mat <- listw2mat(sp_weights)

  # Q - where u = Qe, Q = (I - pW)^-1
  q_mat <- solve(diag(my_model$dims$N) - rho*w_mat)

  # Omega - QQ'
  omega_mat <- q_mat%*%t(q_mat)

  # X
  x_mat <- my_model$my_data %>% 
    dplyr::select(-c(LNP_Percent, year, DivisionNm)) %>% 
    mutate(Intercept = 1) %>% 
    dplyr::select(Intercept, everything()) %>% 
    as.matrix()

  # Beta
  beta_mat <- my_model$coefficients

  # T value
  t = qt(0.975, nrow(my_model$gls_data)-ncol(my_model$gls_data))

  # Lambda matrix (FGLS)
  x <- round(seq(min(as.numeric(x_mat[, varname])), max(as.numeric(x_mat[, varname])), 0.025), 3)
  lambda_mat <- data.frame(matrix(0, nrow = length(x), ncol = ncol(x_mat)))
  names(lambda_mat) <- dimnames(x_mat)[[2]]
  lambda_mat[, varname] <- x
  lambda_mat$Intercept <- 1
  lambda_mat <- as.matrix(lambda_mat)

  # Confidence interval
  plot_df <- data.frame(variable = x, fitted = lambda_mat%*%beta_mat, variance = 0)

  for (i in 1:nrow(lambda_mat)) {
    lambda <- lambda_mat[i, ]
    plot_df$variance[i] = sigma^2 * t(lambda) %*% 
      solve(t(x_mat) %*% solve(omega_mat) %*% x_mat) %*% 
      lambda
  }

  plot_df <- plot_df %>% 
    mutate(upper95 = fitted + t*sqrt(variance), lower95 = fitted - t*sqrt(variance))

  # Partial residuals
  points_df <- data.frame(
    variable = my_model$my_data[, varname] %>% unname,
    part_res = (my_model$my_data$LNP_Percent - x_mat%*%my_model$coefficients) + my_model$coefficients[varname]*x_mat[, varname] + my_model$coefficients[1]
  )

  # Plot
  myplot <- ggplot(data = plot_df) + 
    geom_ribbon(aes(x = variable, ymin = lower95, ymax = upper95), fill = "grey80") + 
    geom_point(aes(x = variable, y = part_res), data = points_df, size = 0.75, col = "grey50") + 
    geom_line(aes(x = variable, y = fitted), col = "blue", size = 1) +
    #geom_hline(aes(yintercept = min(upper95)), col = "red") +
    #geom_hline(aes(yintercept = max(lower95)), col = "blue") +
    theme_bw() + 
    labs(x = varname, y = "Response") + 
    ggtitle(year) +
    theme(plot.title = element_text(face = "bold", size = 10, hjust = 0.5))

  if (nolabs == TRUE) {
    myplot <- myplot + labs(x = "", y = "")
  }

  if (!is.null(xlimits) & !is.null(ylimits)) {
    myplot <- myplot + coord_cartesian(xlim = xlimits, ylim = ylimits)
  }

  return(myplot)
}

# Grid visreg
grid_visreg <- function(varname, xvar_title, xlimits, ylimits) {
  p16 <- my_visreg(glsmod16, sp_weights_16, varname = varname, nolabs = T, year = "2016", xlimits, ylimits)
  p13 <- my_visreg(glsmod13, sp_weights_13, varname = varname, nolabs = T, year = "2013", xlimits, ylimits)
  p10 <- my_visreg(glsmod10, sp_weights_10, varname = varname, nolabs = T, year = "2010", xlimits, ylimits)
  p07 <- my_visreg(glsmod07, sp_weights_07, varname = varname, nolabs = T, year = "2007", xlimits, ylimits)
  p04 <- my_visreg(glsmod04, sp_weights_04, varname = varname, nolabs = T, year = "2004", xlimits, ylimits)
  p01 <- my_visreg(glsmod01, sp_weights_01, varname = varname, nolabs = T, year = "2001", xlimits, ylimits)

  plots <- grid.arrange(p01, p04, p07, p10, p13, p16, nrow = 2, 
    left = grid.text("Two-party preferred vote (%)", gp = gpar(cex = 0.8), rot = 90), 
    bottom = grid.text(xvar_title, gp = gpar(cex = 0.8)))

  return(plots)
}

```

```{r plotincomes, fig.cap = "Effect of Incomes", out.width="100%"}
p <- grid_visreg("Incomes", xvar_title = "Incomes", xlimits = c(-1.5, 3.6), ylimits = c(30, 95))
```

```{r plotunemploy, fig.cap = "Effect of Unemployment", out.width="60%"}
p <- grid_visreg("Unemployment", xvar_title = "Unemployment", xlimits = c(-2.1, 3.4), ylimits = c(28, 68))
```

### Industry and type of work
Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as `Extractive` industries) are consistently linked with higher support for the Liberal party, with the mangitude of this effect slightly increasing over the years (see Figure \@ref(fig:plotextractive)). This is unsurprising, as the Liberal party has close ties with these traditional energy industries, and typically present policies to reduce taxation on energy production. Furthermore, electorates with more workers in construction or manufacturing industries (`Transformative`), as well as wholesale trade, retail trade, transport, post or warehousing (`Distributive`), are also linked with the Liberal party. However, these effects are not as strong, and `Transformative` was marginally insignificant in the 2010 election, whereas `Distributive` was not a key socio-demographic in 2001 or 2016. These effects are shown in Figures \@ref(fig:plottransformative) and \@ref(fig:plotdistributive).

```{r plotextractive, fig.cap = "Effect of Extractive", out.width="60%"}
p <- grid_visreg("Extractive", xvar_title = "Extractive", xlimits = c(-0.7, 3.4), ylimits = c(35, 81))
```

```{r plottransformative, fig.cap = "Effect of Transformative", out.width="60%"}
p <- grid_visreg("Transformative", xvar_title = "Transformative", xlimits = c(-2.25, 2.9), ylimits = c(34, 75))
```
```{r plotdistributive, fig.cap = "Effect of Distributive", out.width="60%"}
p <- grid_visreg("Distributive", xvar_title = "Distributive", xlimits = c(-3.1, 2.5), ylimits = c(34, 65))
```

Similarly, the proportions of workers in managerial, administrative, clerical and sales roles (`ManagerAdminClericalSales`) is also a significant predictor of two-party preference across all six elections, with higher proportions supporting the Liberal party.

```{r plotmanager, fig.cap = "Effect of ManagerAdminClericalSales", out.width="60%"}
p <- grid_visreg("ManagerAdminClericalSales", xvar_title = "ManagerAdminClericalSales", xlimits = c(-3.75, 2.8), ylimits = c(32, 70))
```

### Household mobility
```{r plotdiffaddress, fig.cap = "Effect of DiffAddress", out.width="60%"}
p <- grid_visreg("DiffAddress", xvar_title = "DiffAddress", xlimits = c(-2.1, 3.65), ylimits = c(34, 81))
```

In each of the six elections, electorates with a higher proportion of people that have recently moved house (meaning in last five years) were more likely to support the Liberal party, with this effect becoming stronger since 2010. Having controlled for characteristics of house ownership and rental prices (via the factors `PropertyOwned` and `RentLoan` respectively), this is effect is somewhat surprising. *Add more here?*

### Indigenous population
```{r plotindigenous, fig.cap = "Effect of Indigenous population", out.width="60%"}
p <- grid_visreg("Indigenous", xvar_title = "Indigenous", xlimits = c(-0.6, 9.5), ylimits = c(35, 120))
```

The indigenous population of an electorate is found to be a significant predictor in every election, with higher indigenous populations supporting the Liberal party. There are a few outliers with Lingiari (NT) being the most extreme. However, Lingiari itself has been a strong Labor supporter, and this effect appears to be driven by the other electorates with high Indigenous populations (Kalgoorlie (WA), Leichhardt (NSW), Durack (WA), Parkes(NSW) and Kennedy (QLD)), which have almost always elected a Liberal candidate.

### Relationships and families
In each of the six elections, higher proportions people in de facto marriages are tied to support for the Labor party, whereas the effect of marriages themselves are insignificant. Larger family and household size (via the `FamHouseSize` factor) is also linked with Labor support, although only being significant in the 2007 and 2010 elections (see Figure \@ref(fig:plotfamhousesize)).

```{r plotdefacto, fig.cap = "Effect of DeFacto", out.width="60%", eval = F}
p <- grid_visreg("DeFacto", xvar_title = "DeFacto", xlimits = c(-2.2, 5.1), ylimits = c(12, 83))
```

```{r plotfamhousesize, fig.cap = "Effect of FamHouseSize", out.width="60%", eval = F}
p <- grid_visreg("FamHouseSize", xvar_title = "FamHouseSize", xlimits = c(-3.4, 2.7), ylimits = c(12, 83))
```

### Diversity
In the lead up to the 2001 election, the Liberal government (in power) refused permission for the MV Tampa, a ship carrying 433 rescued refugees and 5 crew to enter Australian waters. This, along with the 9/11 attacks, were two key issues that are widely believed to have pushed the Liberal party to victory. In this 2001 election, electorates with a higher proportion of people speaking languages other than english at home (`OtherLanguageHome`) were significantly more likely to support Labor. This remains negative up until the 2016 election albeit insignificant. Higher proportions of people born in Asia (`Born_Asia`) is also significantly linked with Labor support in 2010, which may be due to refugee policy again arising as a key election issue in lead up to the election.

```{r plototherlanguagehome, fig.cap = "Effect of OtherLanguageHome", out.width="60%", eval = F}
p <- grid_visreg("OtherLanguageHome", xvar_title = "OtherLanguageHome", xlimits = c(-1.1, 3.5), ylimits = c(5, 80))
```

### A note on similar variables

Many of the Census variables represent similar information, which is why factors were created and some variables were removed. However, there still remain some variables which are closely related. For example, `OtherLanguageHome` is the proportion of households that speak a language other than English at home, which is likely to be related to the proportion of people born overseas (a sum of the variables with a `Born` prefix). In 2004, the coefficient estiamte for `OtherLanguageHome` negative but not significant, and all `Born` variables are insignificant. If all `Born` variables are removed from the model, `OtherLanguageHome` absorbs the negative effect, becoming significant ($p = 0.01$).

```{r removeolh, eval = F}
OLH04 <- my_fgls(full_formula, 
  my_data = model_df %>% filter(year == "2004") %>% dplyr::select(-starts_with("Born")),
  sp_weights = sp_weights_04)

summary(OLH04)$tTable
```

## A closer look at the residuals

```{r resids, include = F}
all_resids <- data.frame(
  Residuals = c(glsmod16$actual_residuals, glsmod13$actual_residuals, glsmod10$actual_residuals, glsmod07$actual_residuals, glsmod04$actual_residuals, glsmod01$actual_residuals),
  gls_residuals = c(glsmod16$residuals, glsmod13$residuals, glsmod10$residuals, glsmod07$residuals, glsmod04$residuals, glsmod01$residuals),
  bind_rows(
    glsmod16$my_data %>% left_join(tpp16 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"), 
    glsmod13$my_data %>% left_join(tpp13 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"), 
    glsmod10$my_data %>% left_join(tpp10 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"), 
    glsmod07$my_data %>% left_join(tpp07 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"), 
    glsmod04$my_data %>% left_join(tpp04 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"), 
    glsmod01$my_data %>% left_join(tpp01 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"))
) 
#%>% dplyr::select(Residuals, DivisionNm, UniqueID, StateAb, year, LNP_Percent, gls_residuals)
```

### Residuals by state

It is often hypothesized that states have systematic differences that cause their electorates to vote differently. Boxplots of residuals grouped by state reveal that only Tasmania and the Australian Capital Territory appear to have a state-specific effect that is not captured by the models. These states appear to have a bias towards Labor. There are few electorates in these states (five and two, respectively), so this might be due to incumbent effects rather than an actual state-specific bias.

```{r resstate, fig.cap = "Boxplot of residuals by state.", out.width="75%"}
all_resids %>% 
  ggplot(aes(x = StateAb, y = Residuals, col = StateAb)) + 
  geom_boxplot() + 
  geom_jitter(alpha = 0.4) +
  #geom_jitter(aes(label = DivisionNm, shape = year), alpha = 0.4) +
  #facet_wrap(~year) + 
  guides(col = F, shape = F, DivisionNm = F)
```

### Outlier electorates

In

```{r readcooksd, include = F, echo=FALSE}
read_chunk('cooksd.R')
```

```{r getcooksd}
<<GetCookDistance>>
```

```{r cooksd, include = F}
cooksd_df <- data.frame(
  Residuals = c(glsmod16$actual_residuals, glsmod13$actual_residuals, glsmod10$actual_residuals, glsmod07$actual_residuals, glsmod04$actual_residuals, glsmod01$actual_residuals),
  gls_residuals = c(glsmod16$residuals, glsmod13$residuals, glsmod10$residuals, glsmod07$residuals, glsmod04$residuals, glsmod01$residuals),
  cooksd = c(glsmod16$cooksd, glsmod13$cooksd, glsmod10$cooksd, glsmod07$cooksd, glsmod04$cooksd, glsmod01$cooksd),
  bind_rows(
    glsmod16$my_data %>% left_join(tpp16 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"), 
    glsmod13$my_data %>% left_join(tpp13 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"), 
    glsmod10$my_data %>% left_join(tpp10 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"), 
    glsmod07$my_data %>% left_join(tpp07 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"), 
    glsmod04$my_data %>% left_join(tpp04 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"), 
    glsmod01$my_data %>% left_join(tpp01 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"))
) 
```

```{r plotcooksd, fig.cap = "Electorates' Cook's Distance values in each election."}
cooksd_df %>% 
  ggplot(aes(x = cooksd)) + 
  geom_dotplot() + 
  facet_wrap(~year, nrow = 1) + 
  coord_fixed(ratio = .5) +
  labs(x = "Cook's Distance", y = "")
```

```{r}
cooksd_df %>% 
  filter(cooksd > 0.13) %>% 
  select(DivisionNm, StateAb, year, cooksd, Residuals, LNP_Percent, everything()) %>% View
```

The dotplots in Figure \@ref(fig:plotcooksd) show that there are a few outliers in each election, using cook's distance as a measure of influence. The electorate of Sydney (NSW) has the largest cook's distance in 2001, 2004, 2007 and 2010, due to its diverse population (language, birthplace and religion), high number of defacto relationships,  high income, high household mobility and small amount of workers in extractive, distributive and transformative jobs. It has remained a strong supporter of the Labor party and is severely underpredicted by the model. Nearby in metropolitan NSW, the electorate of Wentworth is found to be an outlier in the 2010, 2013 and 2016 elections - coinciding with its Liberal member, Malcolm Turnbull, being the leader of the Liberal party. Wentworth was historically Liberal but increased its two-party vote by over 10 percentage points in 2010 with Turnbull as Liberal leader, whilst experiencing minimal change in its socio-demographics, causing the models to heavily underpredict its Liberal preference.

Melbourne Ports, an upper-middle class electorate just south of Melbourne's the central business district,

### Electorates with largest residuals

The electorates Capricornia (QLD), Leichhardt (QLD) and Riverina (NSW) have the three largest maximum absolute residuals across the six elections. Capricornia is a swing electorate, which strongly supported Labor between 2001 and 2010, before being narrowly won by Liberal in the last two elections. It has many people working in extractive jobs, an above-average Christian population, and above-average proportion of de facto relationships. The model tends to overpredict its two-party preferred vote (in favour of the Liberal party). Leichhardt is an area encompassing the northernmost tip of Queensland and consists of mainly small indigenous communities and uninhabited rural areas, which was won by the Liberal party in all but the 2007 election. It too has an above-average proportion of de facto relationships, along with an above-average proportion of people moving houses and a high proportion of single parent households. The model consistently underpredicts its two-party vote. Riverina is a rural electorate of NSW, and has been a very safe Liberal (National) seat since 1980. It has a relatively monocultural population consisting of a high proportion of Australian citizens and Christians, an above-average amount of extractive jobs and below average education levels. Its Liberal preference is also underpredicted.

```{r eval = F}
all_resids %>% 
  group_by(UniqueID) %>% 
  mutate(max_res = max(abs(Residuals))) %>% 
  arrange(-max_res) %>% 
  filter(max_res > 13.4) %>% 
  dplyr::select(DivisionNm, LNP_Percent, year, everything())

all_resids %>% 
  group_by(year) %>% 
  mutate(yr_max = max(abs(Residuals))) %>% 
  filter(abs(Residuals) == yr_max) %>% 
  select(DivisionNm, StateAb, year, Residuals, LNP_Percent)

cooksd_df %>% 
  mutate(LNP_win = ifelse(LNP_Percent > 50, 1, 0),
    LNP_predicted = ifelse(LNP_Percent - Residuals > 50, 1, 0),
    correct = ifelse(LNP_win == LNP_predicted, 1, 0)) %>% 
  filter(correct == 0) %>% dim
  group_by(year) %>% 
  mutate(yr_max = max(abs(Residuals))) %>% 
  filter(abs(Residuals) == yr_max) %>% 
  select(DivisionNm, StateAb, year, Residuals, LNP_Percent)
```

```{r eval = F}
library(plotly)
ggplot(all_resids) + geom_jitter(aes(x = year, y = Residuals, label = DivisionNm))
ggplotly()

all_resids %>% 
  group_by(UniqueID) %>% 
  mutate(mean_res = mean(abs(Residuals)), min_res = min(abs(Residuals))) %>%
  filter(min_res > 5) %>%
  ungroup() %>% dplyr::select(Residuals, DivisionNm, year, min_res, mean_res, StateAb) %>% View
  ggplot(aes(x = year, y = Residuals, group = DivisionNm, col = DivisionNm, label = DivisionNm)) + 
  geom_line() + geom_point() +
  guides(col = F)
```