### 2019-03-29 old_book_files/05-conclusion.Rmd

# Conclusion

Across the past six elections, most of the socio-demographics that drive the electoral two-party preferred vote have remained steady, whilst a few (typically weaker) effects have varied over time. Industry and type of work are particularly influential, with energy-related and manufacturing/construction jobs, as well as administrative roles being strongly linked with the Liberal party in all elections. Incomes have had a similarly consistent effect, with higher income areas supporting Liberal. Higher levels of unemployment have shifted from weak association with Labor to a significant Liberal effect over the years, and higher education levels have been associated with Labor since 2007 (although only significant in 2016). It is also found that electorates with higher household mobility support Liberal, birthplace diversity favours Labor and more de facto relationships align with Labor preference - although marriages, family and household sizes have no material influence. Furthermore, the neighbourhood (spatial) effects were found to be positive in all elections, although only significant in 2001 and 2016, meaning that in the 2004-2013 elections, electorates effectively voted independently.

The findings in this paper complement the existing literature by modelling temporal trends, which as far as the authors are aware, has not been done previously for Australian elections using a regression framework. It is also the first study to model any Australian election since 2010 using Census information. Additionally, a key contribution of this research is the spatio-temporal imputation, which is applied to obtain Census data for the 2004, 2007, 2010 and 2013 elections. All data sets used in this study have been contributed to the `eechidna` `R` package, which provides a rich, accessible data resource for future Australian electoral analysis.