### 2019-03-29 old_book_files/07-software.Rmd

# Software

All election and Census data sets, along with electoral maps and more, are available in the `eechidna` `R` package, which can be downloaded from CRAN. `eechidna` (Exploring Election and Census Highly Informative Data Nationally for Australia) makes it easy to look at the data from the Australian Federal elections and Censuses that occurred between 2001 and 2016. This study contributed a large revision to the `eechidna` package, which included the addition of election and Census data for 2001-2010, vote outcomes for polling booths and imputed Census data for election years. For more details on using `eechidna`, please see the articles (vignettes) on the github page (https://ropenscilabs.github.io/eechidna/).

The authors would like to give a sincere thanks to Anthony Ebert, Heike Hofmann, Thomas Lumley, Ben Marwick, Carson Sievert, Mingzhu Sun, Dilini Talagala, Nicholas Tierney, Nathaniel Tomasetti, Earo Wang and Fang Zhou, all of whom have contributed to the `eechidna` package.