### 2019-03-29 old_book_files/raw-files/introduction.Rmd

---
title: "introduction"
author: "Jeremy Forbes"
date: "14/03/2019"
output: html_document
---

Australia has changed in many ways over the last two decades. Rising house prices, country-wide improvements in education, an ageing population, and a decline in religious affiliation, are just a few facets of the country's evolving socio-demographic characteristics[^1]. At the same time, political power has moved back and forth between the two major parties. In the 2007 and 2010 federal elections, the Australian Labor Party (Labor) was victorious, whereas the 2001, 2004, 2013 and 2016 elections were won by the Liberal National coalition (Liberal). The two-party preferred vote, a measure of support between these two parties, fluctuated between 47.3% and 53.5% (in favour of the Liberal party) over this period. This study explores how electoral characteristics relate to two-party preference, and whether their effects have changed over time. Electoral socio-demographics are derived from the Census, and vote counts are obtained from the federal election.

Joining these two data sources is problematic as there in an inherent asynchronicity in the two events, a Census is conducted by the Australian Bureau of Statistics (ABS) every five years, whereas a federal election (conducted by the Australian Electoral Commission (AEC)) usually occurs every three years. The first problem addressed is that of obtaining appropriate Census data for the 2004, 2007, 2010 and 2013 elections - election years in which a Census does not occur. The predominant approach in previous studies is to join voting outcomes to the nearest Census, without accounting for any temporal differences (see @DavisStimson98, @Stimson06, @Liao09 and @Stimson09). Furthermore, electoral boundaries change regularly, so spatial discrepancies also arise when matching electoral data. To obtain appropriate Census data for these four elections, electoral socio-demographics are imputed using a spatio-temporal imputation that combines areal interpolation [@Goodchild1993] and linear time-interpolation. The process of collection and wrangling the raw data is outlined in @CHAPTER2, and the imputation process is detailed in @CHAPTER3. All data and associated documentation relating to this procedure has been made available in the `eechidna` `R` package on `CRAN`, providing a resource for future analysis.

Previous work on modelling Australian federal elections have found that aggregate socio-demographics are relatively good predictors of voting outcomes. @Forrest01 does this using multiple regression of the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998. @Stimson06, @Stimson09 and @Stimson12 use principal component analysis of polling booths in the 2001, 2004 and 2007 elections respectively, also finding that socio-demographic characteristics of polling booths are linked to their two-party preferred vote. On the contrary, @Stimson09 models the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, finding that little of swing vote can be explained by Census data. Instead of analyzing a single election in isolation, this paper employs a consistent model framework across six elections so that temporal changes in the effects of socio-demographics can be observed, where each federal elections is modelled with a cross-sectional data set. The use of a regression framework to examine these socio-political relationships over time is seemingly absent from previous Australian studies. It also appears that no study has attempted any type of statistical analysis of socio-demographics in conjunction with voter behaviour in Australia since 2007, making this paper distinctly different from those previous.

The cross-sectional data set for each election consists of the two-party preferred vote (response variable), and socio-demographic variables (explanatory variables) that characterise each electorate. To obtain these cross-sections, socio-demographic variables are first standardized, and then principal components are used to group variables into "factors". To account for the inherent spatial structure of the data, a spatial error model is fit for each election. These steps are discussed in @CHAPTER4. In @CHAPTER5 inference and diagnostics are conducted on the models to see which effects are significant, how effects change over time and which electorates have abnormal voting behaviour.

---------

, and then selecting the most influential variables with an information-theoretic approach via Akaike weights to determine relative variable importance [@BurnhamAnderson2002]. Spatial error models are then fit to account for spatial dependence.

Previous work on modelling voting outcomes as a function of aggregate socio-demographics is focused on the aggregate level of polling booths - a finer level of aggregation than electorates. @Forrest01 use multiple regression to find that polling booth socio-demographic covariates explained a lot of variation in the primary vote for both Liberal and Labor parties in the Farrer electorate during the 1998 election. @Stimson09 also apply a multiple regression framework for the swing vote (change in two-party preferred vote) in the 2007 election, but resultant model has a relatively poor fit. Principal component analysis has also been used in @Stimson06, @Stimson09 and @Stimson12 to examine the 2001, 2004 and 2007 elections, respectively.

, which find that the Labor party was found to appeal to lower income, younger and more multicultural areas, whereas the Liberal party was associated with higher income, older and more mono-cultural areas in the 2001, 2004 and 2007 elections.

Joining aggregate socio-demographics information from the Census with federal election voting data is problematic, as there in an inherent asynchronicity in the two events - a Census occurs every five years, whereas an election usually occurs every three years. The predominant approach in previous studies is to join voting outcomes to the nearest Census, without accounting for any temporal differences (see @DavisStimson98, @Stimson06, @Liao09 and @Stimson09). Furthermore, electoral boundaries change regularly, so spatial discrepancies also arise when matching electoral data. To combat

Previous studies of aggregate socio-demographics characteristics and voting outcomes commonly use a regression framework and typically use polling booths - rather than electorates - as the level of aggregation, as in @DavisStimson98, @Forrest01, @Stimson06, @Liao09 and @Stimson09. Those that use regression to identify significant determinants of party support find that

@Forrest01 use multiple regression to identify significant determinants of primary vote support in the 1998 Farrer electorate, finding that socio-demographic covariates explained a lot of variation for  Liberal and Labor support. A similar exercise

@Forrest01 use multiple regression to identify significant determinants of the primary vote for each party in the Farrer electorate during the 1998 election

@Stimson09 model the swing vote in the 2007

To obtain data aggregate socio-demographics characteristics are obtained from the Australian Census of Population and Housing and voting data from the Australian Electoral Commission. However, almost all

and voting outcomes involves joining socio-demographic information from the Australian Census of Population and Housing (Census) with voting data from the Australian Electoral Commission. The predominant approach to use polling booth level data,

--------

Few existing studies have conducted an analysis of Australian federal elections using socio-demographic information. Amongst these, the predominant approach is to merge data from the most recent Census that occurs prior to the election of interest. This is done using GIS data obtained at the collection district (Census) and polling booth (election) aggregate levels, as in @DavisStimson98, @Forrest01, @Stimson06, @Liao09 and @Stimson09. Collection districts are matched to the nearest polling booth. Our study is focused on electorates, which reflects a greater level of aggregation than polling booths. Therefore, we adopt a different method of matching Census information with electoral voting outcomes, based on the piece-wise approximation method of areal imputation in @Goodchild1993.

From our revive of the literature, a regression framework using socio-demographics to understand Australian election outcomes has been applied on only two occasions. @Forrest01 use multiple regression to identify significant determinants of the primary vote for each party in the Farrer electorate during the 1998 election, by modelling each party's share of the primary vote as a separate regression, each using ordinary least squares. The adjusted R-squared values obtained from these regressions range from 0.158 to 0.785, suggesting that some parties (including the Labor and Liberal parties) are more predictable than others. They found that age, education and migrant populations were important in relation to support for both Labor and Liberal. @Stimson09 also apply a multiple regression framework to party preference, by modelling the swing vote in the 2007 election. The resultant model has relative poor fit (adjusted R-squared of 0.223), and as such no conclusions about socio-demographic effects were made. Both of these studies use a step-wise procedure to reduce the variable set, and had limited success in providing rich insights into socio-demographic preferences.

Applying a step-wise procedure to a data set with a low number of observations per predictor is problematic [@Steyerberg2009]. In our study, each election is modelled as a separate cross-section containing 63 socio-demographic predictors with only 150 observations. Therefore, instead of step-wise selection, we adopt an information-theoretic approach that uses Akaike weights to determine relative variable importance [@BurnhamAnderson2002], and use this as a criterion for variable selection. Akaike weights are often used in ecological studies (see @Johnson2004 and @Symonds2011).

Another method to differentiate between party preference is to use principal component analysis on the socio-demographic variables, as shown in @Stimson06, @Stimson09 and @Stimson12. Electorates are assigned to a particular party based on their primary vote, and each party is then summarized by its mean z-scores in the directions of the first two principal components. For the 2001, 2004 and 2007 elections, the Labor party was found to appeal to lower income, younger and more multicultural areas, whereas the Liberal party was associated with higher income, older and more mono-cultural areas [@Stimson12]. Whilst we do not assign parties to electorates in this way, we do use principal components to understand and group similar predictor variables together.

Other election-related studies that aimed to understand party preference predominantly modelled individual voter behaviour using survey data (see @Schram92, @Wilson2011 and @Gibson2011).

Seemingly absent from previous Australian studies is the use of a regression framework to examine these socio-political relationships over time. It also appears that no study has attempted any type of statistical analysis of socio-demographics in conjunction with voter behaviour in Australia since 2007. In this thesis, all six elections thus far held in the 21st century are examined.

--------

To understand

Previous studies of Australian federal elections have shown that socio-demographic characteristics (simply, socio-demographics) are good predictors of party support, at an aggregate level. However, these studies have not ventured beyond the 2007 election. Adding to the existing research, this thesis explores the dynamics of electoral party preference for the six federal elections between 2001 and 2016, using a statistical framework. The two-party preferred vote is the outcome of interest, and is modelled as a function of electoral socio-demographics. An emphasis is placed on understanding how the effects of socio-demographics on voting preferences have changed over time.

In the last two decades, political power in Australia has moved back and forth between the two major parties, the Australian Labor Party (Labor) and the Liberal National coalition (Liberal). over the same period Australia has also experienced rising house prices, improvements in education and an ageing population, amongst other changes to its national demographic. These simultaneous changes elicit questions about the relationships between socio-demographics and voting behaviour. Previous work in this area predominantly combine socio-demographic information from the nearest Census, and attach it to the election outcomes at an aggregate level. This poses problems due to temporal differences, and also a spatial mismatch - as the boundaries for aggregate areas (Federal electorates, for example) often change.

This paper outlines an approach to imputing Census information that accounts for both spatial and temporal differences. It then follows with an application of modelling the electoral two-party preferred vote over the past six Australian federal elections.

A key motivation in gathering and imputing this Australian data is to make it easily accessible via the `eechidna` R package, which is available on CRAN.

In the 2007 and 2010 federal elections, the Australian Labor Party (Labor) was victorious, whereas the 2001, 2004, 2013 and 2016 elections were won by the Liberal National coalition (Liberal). A measure of support between these two parties is the two-party preferred vote, which has been as high as 53.5% and as low as 47.3% in favour of the Liberal party over the six elections.

Few studies

The simultaneous changes in characteristics of the Australian population and preference between the two parties lead to our research question: *which socio-demographic characteristics are key determinants of electoral support between the Labor and Liberal parties, and how have their effects changed over time?*

Previous studies of Australian federal elections have shown that socio-demographic characteristics (simply, socio-demographics) are good predictors of party support, at an aggregate level. However, these studies have not ventured beyond the 2007 election. Adding to the existing research, this thesis explores the dynamics of electoral party preference for the six federal elections between 2001 and 2016, using a statistical framework. The two-party preferred vote is the outcome of interest, and is modelled as a function of electoral socio-demographics. An emphasis is placed on understanding how the effects of socio-demographics on voting preferences have changed over time.

The socio-demographic information is obtained from the Australian Census of Population and Housing (Census), a country-wide survey conducted every five years. Federal elections are generally conducted every three years. There are 150 electorates in Australia, with each corresponding to a single seat in the federal House of Representatives. It is electorates that ultimately determine the result of an Australian federal election[^2], which is why electorates are chosen as the level of aggregation for this study. Therefore, the information in this study consists of the electoral two-party preferred vote, and Census data aggregated to Commonwealth Electoral Divisions, which is an approximation of the electoral boundaries.

The two-party preferred vote and the socio-demographic data from the Census possess a temporal component and a spatial component, as both are observed at particular points in time and are aggregate measures for a set of regions in Australia. Censuses and elections often occur in different years, and electoral boundaries change frequently. Therefore, in order to match socio-demographic information from the Census with the electoral two-party preferred vote for each election, a method of spatio-temporal imputation is required. The method used in this study involves estimating an electorate's socio-demographic composition at the time of a Census, using an areal imputation approach, and then linearly interpolating between Census years to approximate the socio-demographics of that electorate, at the time of the election. To adjust for changing variable scales across elections, each socio-demographic variable is centered to have a mean of zero and standardized to have a variance of one, within each election. The resultant cross-sectional data set for each election consists of $n=150$ electorates (observations) and $p=63$ socio-demographic (explanatory) predictors.

Each election is modelled separately with a consistent model specification. This avoids imposing a time-invariant restriction on any regressor, and allows for socio-demographic effects to be compared across elections. In order to obtain the desired model specification, steps are taken to reduce the dimension of the data set, as a large number of predictors $p=63$ and relatively small sample size $n=150$ is problematic.

The first step in dimension reduction involves grouping similar socio-demographic variables together into *factors*. Principal component analysis of socio-demographic variables is employed to determine which variables should be grouped, by identifying variables that have large loadings in one the first four components. Factors are also centered and standardized. The result is a reduced set of $p=30$ predictors. From this reduced set, all possible five variable linear models are fit for each election, with the two-party preferred vote as the response. These models form a set for which Akaike weights are calculated, providing a measure relative variable importance for that election. Using this information-theoretic technique, the five most important variables are identified for each of the six elections. A superset is formed by taking the union of these important socio-demographics, which results in a set of ten variables.

The final specification consists of regressing the two-party preferred vote on the superset of ten predictors and four two-way interactions. Each election model is estimated separately using ordinary least squares. Insights are then found by visualizing the socio-demographic effects within each election. The estimated main effects are described using partial residual plots, and two-way interactions are illustrated with two-dimensional contour plots.

To guide the reader through each step of the analysis, this thesis is structured as follows. Chapter \@ref(ch:litreview) contains a literature review of Australian federal elections modelled using Census information. The data used in this study and the imputation process are described in Chapter \@ref(ch:Data). The steps involved in choosing a statistical model, reducing the dimension of the data, selection of the most important socio-demographic variables, and the final model specification, are explained in Chapter \@ref(ch:Modelling). Chapter \@ref(ch:Insights) contains the insights from the final estimated models. The conclusion and a discussion of possible extensions are detailed in Chapter \@ref(ch:Conclusion).

A key deliverable of this project is the contribution made to the existing `eechidna`[^3] `R` package. This includes the GIS maps, data from all Censuses and elections, and the imputed socio-demographic profiles of each electorate at the time of each election created during this project. When future elections and Censuses come around, the `eechidna` package and the work in this project will provide a resource for anyone wishing to conduct their own socio-political analyses.

[^1]: Socio-demographics are characteristics of a population, such as the distributions of age, gender, ethnicity, education level and income. 
[^2]: The party that wins support from a majority of seats in the House of Representatives forms the government.
[^3]: The name `eechidna` is an acronym for 'Exploring Election and Census Highly Informative Data Nationally for Australia'.