### 2019-04-13 draft-paper.tex
tweaked a few figures and language in the results section

\documentclass[11pt,a4paper,]{article}
\usepackage{lmodern}

\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\hypersetup{
            pdftitle={Spatial modelling of the two-party preferred vote in Australian federal elections: 2001--2016},
            pdfkeywords={federal election, Census, Australia, spatial modelling, imputation,
socio-demographics, electorates, R, eechidna},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{geometry}
\geometry{left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm}
\usepackage[style=authoryear-comp,]{biblatex}
\addbibresource{references.bib}
\usepackage{longtable,booktabs}
% Fix footnotes in tables (requires footnote package)
\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\title{Spatial modelling of the two-party preferred vote in Australian federal
elections: 2001--2016}

%% MONASH STUFF

%% CAPTIONS
\RequirePackage{caption}
\DeclareCaptionStyle{italic}[justification=centering]
 {labelfont={bf},textfont={it},labelsep=colon}
\captionsetup[figure]{style=italic,format=hang,singlelinecheck=true}
\captionsetup[table]{style=italic,format=hang,singlelinecheck=true}

%% FONT
\RequirePackage{bera}
\RequirePackage{mathpazo}

%% HEADERS AND FOOTERS
\RequirePackage{fancyhdr}
\pagestyle{fancy}
\rfoot{\Large\sffamily\raisebox{-0.1cm}{\textbf{\thepage}}}
\makeatletter
\lhead{\textsf{\expandafter{\@title}}}
\makeatother
\rhead{}
\cfoot{}
\setlength{\headheight}{15pt}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}
\fancypagestyle{plain}{%
\fancyhf{} % clear all header and footer fields
\fancyfoot[C]{\sffamily\thepage} % except the center
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}}

%% MATHS
\RequirePackage{bm,amsmath}
\allowdisplaybreaks

%% GRAPHICS
\RequirePackage{graphicx}
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\textfraction}{0.15}
\renewcommand{\floatpagefraction}{0.8}

%\RequirePackage[section]{placeins}

%% SECTION TITLES
\RequirePackage[compact,sf,bf]{titlesec}
\titleformat{\section}[block]
  {\fontsize{15}{17}\bfseries\sffamily}
  {\thesection}
  {0.4em}{}
\titleformat{\subsection}[block]
  {\fontsize{12}{14}\bfseries\sffamily}
  {\thesubsection}
  {0.4em}{}
\titlespacing{\section}{0pt}{*5}{*1}
\titlespacing{\subsection}{0pt}{*2}{*0.2}

%% TITLE PAGE
\def\Date{\number\day}
\def\Month{\ifcase\month\or
 January\or February\or March\or April\or May\or June\or
 July\or August\or September\or October\or November\or December\fi}
\def\Year{\number\year}

\makeatletter
\def\wp#1{\gdef\@wp{#1}}\def\@wp{??/??}
\def\jel#1{\gdef\@jel{#1}}\def\@jel{??}
\def\showjel{{\large\textsf{\textbf{JEL classification:}}~\@jel}}
\def\nojel{\def\showjel{}}
\def\addresses#1{\gdef\@addresses{#1}}\def\@addresses{??}
\def\cover{{\sffamily\setcounter{page}{0}
        \thispagestyle{empty}
        \placefig{2}{1.5}{width=5cm}{monash2}
        \placefig{16.9}{1.5}{width=2.1cm}{MBusSchool}
        \begin{textblock}{4}(16.9,4)ISSN 1440-771X\end{textblock}
        \begin{textblock}{7}(12.7,27.9)\hfill
        \includegraphics[height=0.7cm]{AACSB}~~~
        \includegraphics[height=0.7cm]{EQUIS}~~~
        \includegraphics[height=0.7cm]{AMBA}
        \end{textblock}
        \vspace*{2cm}
        \begin{center}\Large
        Department of Econometrics and Business Statistics\\[.5cm]
        \footnotesize http://monash.edu/business/ebs/research/publications
        \end{center}\vspace{2cm}
        \begin{center}
        \fbox{\parbox{14cm}{\begin{onehalfspace}\centering\Huge\vspace*{0.3cm}
                \textsf{\textbf{\expandafter{\@title}}}\vspace{1cm}\par
                \LARGE\@author\end{onehalfspace}
        }}
        \end{center}
        \vfill
                \begin{center}\Large
                \Month~\Year\\[1cm]
                Working Paper \@wp
        \end{center}\vspace*{2cm}}}
\def\pageone{{\sffamily\setstretch{1}%
        \thispagestyle{empty}%
        \vbox to \textheight{%
        \raggedright\baselineskip=1.2cm
     {\fontsize{24.88}{30}\sffamily\textbf{\expandafter{\@title}}}
        \vspace{2cm}\par
        \hspace{1cm}\parbox{14cm}{\sffamily\large\@addresses}\vspace{1cm}\vfill
        \hspace{1cm}{\large\Date~\Month~\Year}\\[1cm]
        \hspace{1cm}\showjel\vss}}}
\def\blindtitle{{\sffamily
     \thispagestyle{plain}\raggedright\baselineskip=1.2cm
     {\fontsize{24.88}{30}\sffamily\textbf{\expandafter{\@title}}}\vspace{1cm}\par
        }}
\def\titlepage{{\cover\newpage\pageone\newpage\blindtitle}}

\def\blind{\def\titlepage{{\blindtitle}}\let\maketitle\blindtitle}
\def\titlepageonly{\def\titlepage{{\pageone\end{document}}}}
\def\nocover{\def\titlepage{{\pageone\newpage\blindtitle}}\let\maketitle\titlepage}
\let\maketitle\titlepage
\makeatother

%% SPACING
\RequirePackage{setspace}
\spacing{1.5}

%% LINE AND PAGE BREAKING
\sloppy
\clubpenalty = 10000
\widowpenalty = 10000
\brokenpenalty = 10000
\RequirePackage{microtype}

%% PARAGRAPH BREAKS
\setlength{\parskip}{1.4ex}
\setlength{\parindent}{0em}

%% HYPERLINKS
\RequirePackage{xcolor} % Needed for links
\definecolor{darkblue}{rgb}{0,0,.6}
\RequirePackage{url}

\makeatletter
\@ifpackageloaded{hyperref}{}{\RequirePackage{hyperref}}
\makeatother
\hypersetup{
     citecolor=0 0 0,
     breaklinks=true,
     bookmarksopen=true,
     bookmarksnumbered=true,
     linkcolor=darkblue,
     urlcolor=blue,
     citecolor=darkblue,
     colorlinks=true}

%% KEYWORDS
\newenvironment{keywords}{\par\vspace{0.5cm}\noindent{\sffamily\textbf{Keywords:}}}{\vspace{0.25cm}\par\hrule\vspace{0.5cm}\par}

%% ABSTRACT
\renewenvironment{abstract}{\begin{minipage}{\textwidth}\parskip=1.4ex\noindent
\hrule\vspace{0.1cm}\par{\sffamily\textbf{\abstractname}}\newline}
  {\end{minipage}}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\usepackage[showonlyrefs]{mathtools}
\usepackage[no-weekday]{eukdate}

%% BIBLIOGRAPHY

\makeatletter
\@ifpackageloaded{biblatex}{}{\usepackage[style=authoryear-comp, backend=biber, natbib=true]{biblatex}}
\makeatother
\ExecuteBibliographyOptions{bibencoding=utf8,minnames=1,maxnames=3, maxbibnames=99,dashed=false,terseinits=true,giveninits=true,uniquename=false,uniquelist=false,doi=false, isbn=false,url=true,sortcites=false}

\DeclareFieldFormat{url}{\texttt{\url{#1}}}
\DeclareFieldFormat[article]{pages}{#1}
\DeclareFieldFormat[inproceedings]{pages}{\lowercase{pp.}#1}
\DeclareFieldFormat[incollection]{pages}{\lowercase{pp.}#1}
\DeclareFieldFormat[article]{volume}{\mkbibbold{#1}}
\DeclareFieldFormat[article]{number}{\mkbibparens{#1}}
\DeclareFieldFormat[article]{title}{\MakeCapital{#1}}
\DeclareFieldFormat[inproceedings]{title}{#1}
\DeclareFieldFormat{shorthandwidth}{#1}
% No dot before number of articles
\usepackage{xpatch}
\xpatchbibmacro{volume+number+eid}{\setunit*{\adddot}}{}{}{}
% Remove In: for an article.
\renewbibmacro{in:}{%
  \ifentrytype{article}{}{%
  \printtext{\bibstring{in}\intitlepunct}}}

\makeatletter
\DeclareDelimFormat[cbx@textcite]{nameyeardelim}{\addspace}
\makeatother
\renewcommand*{\finalnamedelim}{%
  %\ifnumgreater{\value{liststop}}{2}{\finalandcomma}{}% there really should be no funny Oxford comma business here
  \addspace\&\space}

\jel{C31, C33, D72}

\RequirePackage[absolute,overlay]{textpos}
\setlength{\TPHorizModule}{1cm}
\setlength{\TPVertModule}{1cm}
\def\placefig#1#2#3#4{\begin{textblock}{.1}(#1,#2)\rlap{\includegraphics[#3]{#4}}\end{textblock}}

\nocover

\author{Jeremy~Forbes, Dianne~Cook, Rob J~Hyndman}
\addresses{\textbf{Jeremy Forbes}\newline
Monash University\newline Melbourne, Australia
\newline{Email: \href{mailto:jeremy.forbes@monash.edu}{\nolinkurl{jeremy.forbes@monash.edu}}}\newline Corresponding author\\[1cm]
\textbf{Dianne Cook}\newline
Monash University\newline Melbourne, Australia
\\[1cm]
\textbf{Rob J Hyndman}\newline
Monash University\newline Melbourne, Australia
\\[1cm]
}

\date{\sf\Date~\Month~\Year}
\makeatletter
 \lfoot{\sf Forbes, Cook, Hyndman: \@date}
\makeatother

%% Any special functions or other packages can be loaded here.
\usepackage{float}
\let\origfigure\figure
\let\endorigfigure\endfigure
\renewenvironment{figure}[1][2] {
    \expandafter\origfigure\expandafter[htbp]
} {
    \endorigfigure
}

\let\origtable\table
\let\endorigtable\endtable
\renewenvironment{table}[1][2] {
    \expandafter\origtable\expandafter[htbp]
} {
    \endorigtable
}

% Don't number subsubsections

\titleformat{\subsubsection}
  {\normalfont\bfseries\sffamily}   % The style of the section title
  {}                             % a prefix
  {0pt}                          % How much space exists between the prefix and the title
  {}    % How the section is represented

\begin{document}
\maketitle
\begin{abstract}
We examine the relationships between electoral socio-demographic
characteristics and two-party preference in the six Australian federal
elections held between 2001 to 2016. Socio-demographic information is
derived from the Australian Census, which occurs every five years. Since
a Census is not directly available for each election, spatio-temporal
imputation is employed to estimate Census data for the electorates at
the time of each election. This accounts for both spatial and temporal
changes in electoral characteristics between Censuses. To capture any
spatial heterogeneity, a spatial error model is estimated for each
election, which incorporates a spatially structured random effect vector
that can be thought of as the unobserved political climate in each
electorate. Over time, the impact of most socio-demographic
characteristics that affect electoral two-party preference do not vary,
with industry of work, incomes, household mobility and de facto
relationships having strong effects in each of the six elections.
Education and unemployment are amongst those that have varying effects.
It is also found that between 2004 and 2013, the spatial effect was
insignificant, meaning that electorates voted effectively independently.
All data featured in this study has been contributed to the
\texttt{eechidna} \texttt{R} package (available on CRAN).
\end{abstract}
\begin{keywords}
federal election, Census, Australia, spatial modelling, imputation,
socio-demographics, electorates, R, eechidna
\end{keywords}

\section{Introduction}\label{intro}

Australia has changed in many ways over the last two decades. Rising
house prices, country-wide improvements in education, an ageing
population, and a decline in religious affiliation, are just a few
facets of the country's evolving socio-demographic characteristics. At
the same time, political power has moved back and forth between the two
major parties. In the 2007 and 2010 federal elections, the Australian
Labor Party (Labor) was victorious, whereas the 2001, 2004, 2013 and
2016 elections were won by the Liberal National coalition (Liberal). The
two-party preferred vote, a measure of support between these two
parties, fluctuated between 47.3\% and 53.5\% (in favour of the Liberal
party) over this period. This study explores how electoral
socio-demographic characteristics relate to two-party preference, and
whether their effects have changed over time. Electoral
socio-demographics are derived from the Australian Census, and vote
counts are obtained from Australian federal elections.

Joining these two data sources is problematic as there in an inherent
asynchronicity in the two types of events. A Census is conducted by the
Australian Bureau of Statistics (ABS) every five years, whereas federal
elections, conducted by the Australian Electoral Commission (AEC),
usually occur every three years or so. The first problem addressed is
that of constructing appropriate Census data for the 2004, 2007, 2010
and 2013 elections --- election years in which a Census does not occur.
The predominant approach in previous studies was to join voting outcomes
to the nearest Census, without accounting for any temporal differences
\autocites[see][]{DavisStimson98}{Stimson06}{Liao09}{Stimson09}.
Furthermore, electoral boundaries change regularly, so spatial
discrepancies also arise when matching with electoral data. To obtain
appropriate ``Census-like'' data for these four elections, electoral
socio-demographics are constructed using a spatio-temporal imputation
that combines areal interpolation \autocite{Goodchild1993} and linear
time-interpolation. Collecting and wrangling the raw data, along with
the imputation process, are detailed in Section \ref{data}. All data and
associated documentation relating to this procedure are available in the
\texttt{eechidna} \texttt{R} package \autocite{eechidna}, providing a
resource for future analysis.

Previous work on modelling Australian federal elections has found that
aggregate socio-demographics are relatively good predictors of voting
outcomes. \textcite{Forrest01} used multiple regression to model the
Liberal and Labor primary vote for polling booths in the Farrer
electorate in 1998 as a function of Census variables from 1996.
\textcite{Stimson06}, \textcite{Stimson09} and \textcite{Stimson12} used
principal component analysis of polling booths in the 2001, 2004 and
2007 elections respectively, also finding that socio-demographic
characteristics of polling booths are linked to their two-party
preferred vote. In contrast, \textcite{Stimson09} models the polling
booth swing vote (change in the two-party preferred vote) in the 2007
election, finding that little of the swing vote can be explained by
Census data. Instead of analyzing a single election in isolation, this
paper employs a consistent model framework across six elections so that
temporal changes in the effects of socio-demographics can be observed,
where each federal elections is modelled with a cross-sectional dataset.
The use of a regression framework to examine these socio-political
relationships over time is seemingly absent from previous Australian
studies. It also appears that no study has attempted any type of
statistical analysis of socio-demographics in conjunction with voter
behaviour in Australia since 2007, making this paper distinctly
different from previous work.

The cross-sectional dataset for each election used here consists of the
two-party preferred vote (as the response variable), and a set of common
socio-demographic variables (as the explanatory variables) that
characterize each electorate. To prepare these datasets,
socio-demographic variables are first standardized, and then a principal
component analysis is used to group variables into ``factors''. To
account for the inherent spatial structure of the data, a spatial error
model is then estimated for each election. These steps are discussed in
Section \ref{modelling}. In Section \ref{results} inference is conducted
to see which effects are significant and how these effects change over
time.

Section \ref{conclusion} concludes, followed by two supplementary
sections.

\section{Data collection, wrangling and imputation}\label{data}

\subsection{Collecting the data}\label{collecting-the-data}

The voting outcome of interest is the electoral two-party preferred
vote, which is provided by the Australian Electoral Commission (AEC) for
the 2001, 2004, 2007, 2010, 2013 and 2016 elections via the AEC Tally
Room. The AEC divides Australia into 150 regions, called electorates,
with each corresponding to a single seat in the House of
Representatives. Voting is compulsory in Australia, and each voter
assigns a numbered preference to each available candidate in their
electorate. The two-party preferred vote is determined by a tally of
these preferences where, by convention, only the ranks of the Labor and
Liberal candidates are considered. This is recorded as a percentage
preference in favour of the Liberal party.

Socio-demographic variables were derived from the Australian Census of
Population and Housing (Census), which is a survey of every household in
Australia, recording information such as age, gender, ethnicity,
education level and income. There have been four Censuses so far in the
21st century, conducted in 2001, 2006, 2011 and 2016. The Australian
Bureau of Statistics (ABS) conducts the Census and publishes aggregated
information. The ABS uses electoral boundaries as defined by the AEC at
the time of each Census, which may not match those in place at the
subsequent and previous elections. From the available Census information
aggregated at the electorate level, 65 socio-demographic variables were
defined for each of the electorates to be used in the analysis.

Raw data was sourced online from the AEC and ABS websites in
\texttt{.csv} and \texttt{.xlsx} files. The formats of these files
differ over the years, making extracting the appropriate information a
big task. The functions available in the \texttt{dplyr} \autocite{dplyr}
and \texttt{readxl} \autocite{readxl} \texttt{R} packages are
particularly useful, as they provide fast consistent tools for data
manipulation and functions to import \texttt{.xlsx} files
(respectively). The 2001 and 2006 Census data are published in a format
where the information for each electorate is held in a separate document
making it difficult to use the \texttt{dplyr} tools. Instead, cells have
to be selected from each individual file to construct the desired
variables. All scripts required for the data wrangling process can be
found in the github repository for the \texttt{eechidna} \texttt{R}
package \autocite{eechidna}, along with the raw data. The
\texttt{eechidna} package makes this study entirely reproducible and
provides a resource to help wrangle data for future Censuses and
elections, when they become available.

\subsection{Joining Census and election
data}\label{joining-census-and-election-data}

\subsubsection{Differences between Census and election
data}\label{differences-between-census-and-election-data}

Between 2001 and 2016 there were six elections and four Censuses (see
Figure \ref{fig:timeline}). Electoral boundaries are redistributed
regularly by the AEC, meaning that only in the years where both a Census
and election occur are all boundaries likely to match --- the case for
the 2001 and 2016 elections. Therefore, for the four elections between
2004 and 2013, both temporal and spatial differences in electorates need
to be accounted for when joining the electoral two-party preferred vote
with Census data. For these elections a spatio-temporal imputation
method was employed to obtain electoral socio-demographics. This method
uses Census information from both before and after the election of
interest.

\begin{figure}[h]

{\centering \includegraphics[width=0.9\linewidth]{draft-paper_files/figure-latex/timeline-1}

}

\caption{Timeline of Australian elections and Censuses. They do not always occur in the same year.}\label{fig:timeline}
\end{figure}

\subsubsection{Spatio-temporal
imputation}\label{spatio-temporal-imputation}

For each election, neighbouring Census information has to be combined in
some way so that it represents the boundaries in place at the time of
the election. This is done by taking the electoral boundaries and
imputing the corresponding socio-demographic characteristics for each of
neighbouring Censuses, thereby addressing the spatial aspect. Next, to
deal with the temporal component, characteristics at the time of the
election are constructed using linear interpolation between the
spatially imputed neighbouring Census variables.

To account for spatial differences, the piecewise approximation method
in \textcite{Goodchild1993} is adopted. Consider a map of source zones
\(s = 1,\dots,S\), for which socio-demographic information is available,
and a set of target zones \(t = 1,\dots,T\) for which information is to
be imputed. This is described in the context of a single election, and a
single neighbouring Census.

Let the map of electoral boundaries at the time of a Census define the
source zones, and let the boundaries at the time of the election be the
target zones. Denote the area of intersection between a source zone
\(s\) and a target zone \(t\) as \(A_{s,t}\). Additionally, let the
population of the source zone \(s\) be \(U_s\) and the population of
intersection between source zone \(s\) and target zone \(t\) be
\(P_{s,t}\). The estimated population of intersection is given by \[
  \hat{P}_{s,t} = \frac{U_s*A_{s,t}}{\sum_{t=1}^T A_{s,t}},
  \quad\text{for all $s=1,\dots,S$ and $t=1,\dots,T$}.
\] Note that this estimator implicitly assumes that populations are
uniformly distributed within each source zone.

In order to calculate socio-demographic information for each of the
target zones, a weighted average is taken using the estimated
intersection populations as weights. Denote a given Census variable for
the target zone by \(C_t\), and the same Census variable for the source
zone as \(D_s\). Then, estimate \(C_t\) using \[
  \hat{C}_t = \frac{\sum_{s=1}^{S}{D_s*\hat{P}_{s,t}}}%
  {\sum_{s=1}^{S}{\hat{P}_{s,t}}},
  \quad\text{for each $t=1,\dots,T$}.
\] This concludes the spatial imputation of the socio-demographic
characteristics for one target zone (a single electoral boundary), at
the time of only one of the neighbouring Censuses. This process is
repeated for all of the target zones, and then for the other
neighbouring Census.

\enlargethispage*{0.5cm}

To account for temporal changes, linear interpolation is used between
Census years to get the final estimate of a Census variable for the
target zone in the election year. Let \(y_1\) be the year of the Census
preceding an election, let \(y_2\) be the year of the election, and
\(y_3\) be the year of the Census that follows. Add this year subscript
to the Census variable estimate \(\hat{C}_t\), resulting in
\(\hat{C}_{t,y}\). Linear interpolating between these Census years
results an imputed value for the election year, given by \[
  \hat{C}_{t,y_2} = \frac{y_3-y_2}{y_3-y_1} * \hat{C}_{t,y_1} +
    \frac{y_2-y_1}{y_3-y_1} * \hat{C}_{t,y_3}.
\] Implicitly this assumes that population characteristics change in a
linear manner over time.

\subsubsection*{An illustration of the spatio-temporal
imputation}\label{an-illustration-of-the-spatio-temporal-imputation}
\addcontentsline{toc}{subsubsection}{An illustration of the
spatio-temporal imputation}

Census data is publicly available at different levels of aggregation,
ranging from SA1 (over 50,000 zones) to electoral divisions (150 zones).
For this study, electoral divisions are used as source zones, and the
imputation method is applied to produce socio-demographic variables for
each of the 2004, 2007, 2010 and 2013 elections. As mentioned earlier,
there is no need to impute socio-demographic variables for the 2001 and
2016 elections. To illustrate the method, consider the imputation of
socio-demographic variables for the electorate of Hume in New South
Wales (NSW) at the time of the 2013 federal election. The boundaries
shown in Figure \ref{fig:hume13} define all target zones in NSW for
2013, with the target zone of interest (Hume) shaded purple.

\begin{figure}[h]

{\centering \includegraphics[height=0.28\textheight]{draft-paper_files/figure-latex/hume13-1}

}

\caption{Some of the electoral boundaries in NSW for 2013, with the electoral boundary for Hume shown in purple.}\label{fig:hume13}
\end{figure}

\begin{figure}[h]

{\centering \includegraphics[height=0.28\textheight]{draft-paper_files/figure-latex/hume16-1}

}

\caption{Census division boundaries in NSW for 2016, with the 2013 electoral boundary for Hume, shown in purple. The purple region is not contained within a single Census division.}\label{fig:hume16}
\end{figure}

The corresponding source zones from the 2016 Census are shown in Figure
\ref{fig:hume16}. As can be seen, the Hume boundary from the 2013
election (shaded purple) does not perfectly match any of the source
zones.

There are many source zones from the 2016 Census that intersect with
this purple region, including the divisions of Riverina, Eden-Monaro and
Hume, along with smaller intersecting areas with Fenner, Calare, Gilmore
and Whitlam. The proportion of each source zone that overlaps with the
purple region is calculated, and used to obtain the intersecting
populations \(\hat{P}_{s,t}\).

\begin{table}[H]
\centering\begingroup\fontsize{10}{12}\selectfont

\resizebox{\linewidth}{!}{
\begin{tabular}{lrrrr}
\toprule
Source zone (2016) & Proportion & Source zone population & Population allocated to purple region: $\hat{P}_{s,t}$ & AusCitizen (\%): $D_s$\\
\midrule
Hume & 0.9654 & 150643 & 145427 & 90.0168\\
Riverina & 0.2511 & 155793 & 39117 & 89.1144\\
Eden-Monaro & 0.1109 & 147532 & 16358 & 87.9999\\
Canberra & 0.0028 & 196037 & 548 & 85.4793\\
Fenner & 0.0023 & 202955 & 474 & 83.6432\\
\addlinespace
Whitlam & 0.0006 & 152280 & 92 & 89.5173\\
Gilmore & 0.0006 & 150436 & 86 & 89.0266\\
Calare & 0.0001 & 161298 & 21 & 87.5603\\
\bottomrule
\end{tabular}}
\endgroup{}
\end{table}

Now consider the socio-demographic variable \emph{AusCitizen}, the
proportion of people in the region who are Australian citizens. A
weighted average of \emph{AusCitizen}, with the allocated population
from each source zone as weights, yields
\(\hat{C}_{\text{Hume},2016} = 89.65\%\). Repeating this process using
the 2011 Census yields \(\hat{C}_{\text{Hume},2011} = 91.00\%\).
Finally, linear interpolation between 2011 and 2016 yields the 2013
estimate: \[
  \hat{C}_{\text{Hume},2013}
    = \frac{3}{5} \hat{C}_{\text{Hume},2011} + \frac{2}{5} \hat{C}_{\text{Hume},2016}
    % = \frac{3}{5} \cdot 91.00 \% + \frac{2}{5} \cdot 89.65 \% \\
    =  90.46\%.
\] This is done for each of the 65 socio-demographic variables, and is
repeated for each of the 149 remaining target zones corresponding with
2013 electorates.

\section{Modelling}\label{modelling}

Following this process, electoral socio-demographic variables are
available for each of the six elections and can be joined with their
corresponding two-party preferred votes. Before choosing an appropriate
model, two issues with the socio-demographic variables need to be
addressed. First, variable scales change over the years, making it
important to standardize variables. Second, many variables represent
similar information and where appropriate, should be combined in some
way. Principal component analysis is used to identify variables covary,
from which intuitive groupings are selected to be combined into a single
variable. This also reduces the dimension of the data. After these
steps, a model specification is chosen.

\subsection{Standardizing variables}\label{standardizing-variables}

Many of the socio-demographic variables have changing scales over the
years. For example, inflation-adjusted median rental prices increased
across almost all electorates, with median rent of 200 dollars per week
placing an electorate in the 90th percentile in 2001, but only the 30th
percentile in 2016. In order for socio-demographic effects to be
comparable across years, all explanatory variables are standardized to
have mean zero and variance one within each election year. By
standardizing, each variable is reported as a relative measure compared
to all other electorates in the same year.

\subsection{Creating factors}\label{creating-factors}

There are only \(N = 150\) observations (electorates) in each election
and \(p = 65\) socio-demographic variables in each cross-section, with
many variables represent similar information about an electorate. Any
model that uses all variables would face serious problems with
multi-collinearity and over-fitting, which would likely lead to
erroneous conclusions regarding variable significance. To address this,
groups of variables that represent similar information are combined into
``factors''\footnote{A preliminary step involved removing all age bands,
  because age is represented by median age, and to remove variables
  relating to particular denominations of Christianity.}.

A factor is created from a group of variables if there is an intuitive
reason as to why they should represent similar information and if there
is evidence to suggest that they covary. For example, a potential group
would be variables relating to electoral incomes --- median family,
household and personal incomes. To determine which variables covary,
principal component analysis is used on a combined dataset of
socio-demographic variables from all six elections. It is appropriate to
compute principal components in this way because when computed
separately for each election, scree plots level off after four
components and the loadings of the first four components are similar
across the elections.

Only the first four principal components from the combined dataset are
considered, as the scree plot corresponding to the combined dataset
levels off after the fourth component. Variables that have a large
loading in a particular component are deemed to covary, with a loading
with magnitude greater than 0.15 being considered large. Six factors are
created using this criteria. These are: \texttt{Incomes} (median
personal, household and family incomes); \texttt{Unemployment}
(unemployment and labor force participation rates);
\texttt{ProportyOwned} (rates of housing ownership, mortgages, renting
and government housing); \texttt{RentLoanPrice} (median rental and loan
repayments); \texttt{FamHouseSize} (average household size, ratio of
people to families and household makeup (single person, couple with kids
and couple without kids); and \texttt{Education} (high school and
university qualifications, jobs requiring higher levels of education as
well as vocational course completions and jobs that do not require
higher education levels, such as laborer or tradesperson). For each of
these groups, variables with positive loadings are added and those with
negative loadings are subtracted to create a factor. After computing
these sums, each factor is standardized to have mean zero and variance
one, within each election.

There are \(p=30\) variables in the resultant predictor set, with all of
these used in the regression for each election.

\subsection{Regression incorporating spatially dependent
errors}\label{regression-incorporating-spatially-dependent-errors}

An identical model specification is used for each of the six elections,
with each election modelled separately. This allows the
socio-demographic effects to be estimated separately for each election
year, facilitating analysis of temporal changes in variable effects.
This approach is preferable to using a single longitudinal model because
it avoids any concerns of undue bias stemming from incorrectly imposed
time-varying restrictions on any variable. Without such restrictions, a
pooled cross-sectional model does not yield any distinct advantage over
separate cross-sections. The panel approach is avoided because of how
frequently electoral boundaries change, noting that electorates with the
same name across elections are not guaranteed to represent the same
geographical region. Therefore any fixed or random effects models would
be difficult to estimate without implementing consistent boundaries,
which would require further imputation.

For each cross-section, let the response \(\bm{y}\) be the vector
two-party preferred vote in favour of the Liberal party; for example,
\(y_i = 70\) represents a 70\% preference for Liberal, 30\% for Labor,
in electorate \(i\). Although \(y_i\) lies in the interval \((0,100)\),
observed values are never close to 0 or 100 (minimum \(24.05 \%\) and
maximum \(74.90 \%\)), so there is no need to formally impose the
constraint of \(y_i \in [0,100]\). Furthermore, the responses are found
to be spatially correlated in each election (Moran's I test,
\(p \le 7\cdot10^{-15}\)). This is not surprising as electorates are
aggregate spatial units, and hence the spatial structure of the data
must modelled appropriately.

The spatial error model \autocite{Anselin88} is chosen because it
captures spatial heterogeneity by incorporating a spatially structured
random effect vector \autocite{LeSage2009}. In this context, the random
effect can be thought of as capturing the unobserved political climate
in each electorate, where this climate is correlated with the climate in
neighbouring electorates, under the assumption that the climate is
independent of electoral socio-demographics.

\begin{table}[!htbp] \centering    \caption{Estimated spatial error model parameters (standard errors) for each of the six election years.}    \label{tab:printmodels}  \scriptsize  \begin{tabular}{@{\extracolsep{1pt}}lcccccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]  \\[-1.8ex] & \multicolumn{6}{c}{\vspace{-0.5cm}} \\   & 2001 & 2004 & 2007 & 2010 & 2013 & 2016 \\  \hline \\[-1.8ex]  $\rho$ & 0.46$^{***}$ & 0.29$^{*}$ & 0.24 & 0.19 & 0.27$^{*}$ & 0.50$^{***}$\\  & (0.15) & (0.17) & (0.17) & (0.16) & (0.16) & (0.17)\\  AusCitizen & $-$3.13 & $-$2.64 & $-$2.53 & $-$0.08 & $-$3.40 & $-$1.80 \\    & (2.26) & (2.43) & (2.34) & (2.79) & (2.76) & (2.71) \\    BornAsia & 2.22 & $-$0.95 & $-$1.60 & $-$6.83$^{**}$ & $-$3.03 & $-$0.55 \\    & (2.18) & (2.44) & (2.19) & (2.73) & (2.71) & (2.17) \\    Born\_MidEast & $-$1.15 & $-$1.59 & $-$2.01$^{*}$ & $-$2.03 & $-$0.92 & $-$1.44 \\    & (1.07) & (1.20) & (1.11) & (1.27) & (1.24) & (1.13) \\    BornSEEuro & $-$3.21$^{**}$ & $-$4.24$^{***}$ & $-$3.61$^{***}$ & $-$4.14$^{***}$ & $-$3.69$^{***}$ & $-$2.72$^{***}$ \\    & (1.42) & (1.46) & (1.02) & (1.19) & (1.07) & (0.97) \\    Born\_UK & 0.25 & $-$0.07 & 0.34 & 0.56 & $-$0.09 & $-$1.32 \\    & (1.00) & (0.98) & (0.90) & (1.07) & (1.04) & (1.04) \\    BornElsewhere & $-$5.04 & $-$4.91 & $-$4.13 & 2.35 & $-$5.23 & $-$4.14 \\    & (3.30) & (3.68) & (3.38) & (4.23) & (4.15) & (3.97) \\    Buddhism & $-$0.49 & $-$0.17 & $-$1.37 & $-$0.83 & $-$0.12 & $-$1.60 \\    & (1.39) & (1.61) & (1.61) & (1.80) & (1.68) & (1.56) \\    Christianity & $-$2.48 & $-$1.23 & 0.38 & 0.50 & 2.41 & 1.68 \\    & (1.73) & (1.85) & (1.83) & (1.99) & (1.85) & (1.78) \\    CurrentlyStudying & $-$2.19$^{**}$ & $-$0.13 & 2.06$^{*}$ & 2.12$^{*}$ & 1.15 & $-$0.16 \\    & (0.99) & (1.13) & (1.17) & (1.25) & (1.26) & (1.18) \\    DeFacto & $-$6.44$^{***}$ & $-$5.37$^{**}$ & $-$6.43$^{***}$ & $-$8.07$^{***}$ & $-$6.56$^{**}$ & $-$8.53$^{***}$ \\    & (1.87) & (2.48) & (2.31) & (3.06) & (3.11) & (2.83) \\    DiffAddress & 3.88$^{***}$ & 5.06$^{***}$ & 4.22$^{***}$ & 5.57$^{***}$ & 3.53$^{*}$ & 5.67$^{***}$ \\    & (0.94) & (1.12) & (0.99) & (1.76) & (1.91) & (1.60) \\    Distributive & 1.27 & 2.01$^{*}$ & 1.36 & 1.57 & 2.10$^{*}$ & 1.20 \\    & (1.12) & (1.21) & (1.13) & (1.34) & (1.27) & (1.21) \\    Education & 1.08 & 0.52 & $-$5.52$^{*}$ & $-$4.08 & $-$4.44 & $-$7.07$^{**}$ \\    & (2.38) & (3.12) & (3.27) & (3.95) & (3.78) & (3.55) \\    Extractive & 4.83$^{***}$ & 5.45$^{***}$ & 5.37$^{***}$ & 7.31$^{***}$ & 6.71$^{***}$ & 7.43$^{***}$ \\    & (1.48) & (1.42) & (1.36) & (1.56) & (1.47) & (1.39) \\    FamHouseSize & $-$0.16 & 0.87 & $-$2.40 & $-$2.53 & $-$3.26 & $-$2.91 \\    & (2.19) & (2.72) & (2.69) & (3.25) & (3.28) & (2.90) \\    Incomes & 4.36$^{**}$ & 5.03$^{*}$ & 9.45$^{***}$ & 7.09$^{**}$ & 7.97$^{***}$ & 12.20$^{***}$ \\    & (1.77) & (2.66) & (2.75) & (3.25) & (2.92) & (2.75) \\    Indigenous & 2.91$^{*}$ & 1.97 & 2.48 & 2.84 & 0.67 & $-$0.05 \\    & (1.68) & (1.95) & (1.75) & (2.16) & (2.14) & (2.00) \\    Islam & $-$0.92 & $-$0.97 & $-$0.54 & $-$2.50 & $-$0.82 & $-$0.95 \\    & (1.22) & (1.36) & (1.27) & (1.52) & (1.42) & (1.34) \\    Judaism & 1.88$^{*}$ & 1.78 & 2.66$^{***}$ & 1.97$^{*}$ & 2.74$^{**}$ & 1.65$^{*}$ \\    & (1.05) & (1.13) & (1.01) & (1.15) & (1.10) & (1.00) \\    ManagerAdmin & 2.06$^{***}$ & 3.32$^{***}$ & 6.00$^{***}$ & 5.47$^{***}$ & 5.04$^{***}$ & 5.78$^{***}$ \\    & (0.71) & (0.93) & (0.90) & (1.08) & (1.03) & (1.06) \\    Married & 0.44 & 0.11 & $-$1.22 & $-$0.22 & 0.91 & $-$2.34 \\    & (2.31) & (2.96) & (2.83) & (3.15) & (3.03) & (2.81) \\    MedianAge & 2.32$^{*}$ & 4.96$^{***}$ & 3.66$^{**}$ & 4.00$^{*}$ & 2.30 & 2.87 \\    & (1.32) & (1.65) & (1.81) & (2.26) & (2.08) & (1.79) \\    NoReligion & $-$1.57 & $-$0.92 & 0.56 & $-$0.30 & 1.02 & 1.31 \\    & (1.59) & (1.71) & (1.73) & (1.92) & (1.94) & (2.04) \\    OneParentHouse & $-$1.73 & $-$0.45 & $-$0.75 & $-$1.46 & $-$0.77 & $-$0.74 \\    & (1.36) & (1.59) & (1.49) & (1.69) & (1.57) & (1.47) \\    OtherLanguage & $-$0.44 & 5.92 & 9.98$^{**}$ & 11.24$^{**}$ & 9.00$^{*}$ & 9.84$^{**}$ \\    & (3.22) & (4.16) & (3.91) & (4.76) & (4.66) & (4.44) \\    PropertyOwned & $-$0.46 & $-$0.53 & 0.67 & $-$0.94 & $-$0.48 & 1.41 \\    & (1.37) & (1.50) & (1.43) & (1.76) & (1.67) & (1.50) \\    RentLoanPrice & $-$1.57 & $-$3.32$^{*}$ & $-$4.01$^{**}$ & $-$0.97 & $-$0.70 & $-$0.89 \\    & (1.49) & (1.76) & (1.67) & (2.07) & (2.07) & (2.16) \\    SocialServ & 2.51$^{*}$ & 1.65 & 2.47$^{*}$ & 2.53$^{*}$ & 2.35$^{*}$ & 4.45$^{***}$ \\    & (1.33) & (1.41) & (1.29) & (1.47) & (1.32) & (1.19) \\    Transformative & 3.24$^{**}$ & 4.73$^{***}$ & 4.84$^{***}$ & 4.46$^{**}$ & 3.56$^{**}$ & 4.58$^{***}$ \\    & (1.55) & (1.78) & (1.74) & (1.98) & (1.78) & (1.53) \\    Unemployment & $-$2.45$^{*}$ & $-$3.07$^{*}$ & 0.29 & 0.08 & 1.67 & 2.79$^{**}$ \\    & (1.40) & (1.63) & (1.51) & (1.76) & (1.51) & (1.37) \\    Constant & 50.81$^{***}$ & 52.60$^{***}$ & 47.31$^{***}$ & 49.93$^{***}$ & 53.51$^{***}$ & 50.49$^{***}$ \\    & (0.71) & (0.58) & (0.52) & (0.57) & (0.58) & (0.80) \\   \hline \\[-1.8ex]  Observations & 150 & 150 & 150 & 150 & 150 & 150 \\  Residual Standard Error (GLS) & 4.69 & 5.04 & 4.79 & 5.63 & 5.18 & 4.88\\ \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}  \end{table}

Spatial weights are calculated in accordance with the assumption that an
electorate is equally correlated with any electorate that shares a part
of its boundary. Let \(\rho\) be the spatial autoregressive coefficient,
\(\bm{v}\) be a spherical error term, \(\bm{W}\) be a matrix of spatial
weights (containing information about the neighbouring regions),
\(\bm{X}\) be a matrix of socio-demographic covariates, \(\bm{\beta}\)
be a vector of regression coefficients and \(\bm{a}\) be a spatially
structured random effect vector. \[
  \bm{y} = \bm{X} \bm{\beta} + \bm{a},
\] and \[
  \bm{a} = \rho \bm{W} \bm{a} + \bm{v},
\] where \(\bm{v} \sim N(\bm{0}, \sigma^2 \bm{I_n})\), and hence \[
  \bm{y} = \bm{X} \bm{\beta} + (\bm{I}_n-\rho \bm{W})^{-1}\bm{v}.
\] Estimation of the above spatial error model is undertaken using
feasible generalized least squares.

Table \ref{tab:printmodels} details the estimated model coefficients and
their estimated standard errors, for each of the six elections. An
interpretation of these estimated values is provided in the next
section.

\section{Results}\label{results}

\subsection{Spatial autoregressive
parameter}\label{spatial-autoregressive-parameter}

The spatial autoregressive coefficient \(\rho\) is positive and
significant in only the 2001 and 2016 elections (Figure
\ref{fig:rhovis}), meaning that in these elections, the political
climate of an electorate appears to be affected by the attitudes of its
neighbours. Conversely, in the other four elections, the spatial effect
weakens to become insignificant. In these years, it appears that the
spatial component does not explain anything not already explained by the
electoral socio-demographics, meaning electorates effectively voted
independently.

\begin{figure}[h]

{\centering \includegraphics[height=0.2\textheight,]{draft-paper_files/figure-latex/rhovis-1}

}

\caption{Estimates of the spatial autoregressive parameter for each of the six elections, reported with their individual 95\% confidence intervals. Only in 2001 and 2016 is there a significant spatial component.}\label{fig:rhovis}
\end{figure}

\subsection{Country-wide trend}\label{country-wide-trend}

\begin{figure}[h]

{\centering \includegraphics[height=0.2\textheight,]{draft-paper_files/figure-latex/plotintercept-1}

}

\caption{Estimated intercept for each election, which represents the two-party preferred vote for an electorate with mean characteristics.}\label{fig:plotintercept}
\end{figure}

Since all socio-demographics have been standardized to have a mean of
zero and a variance of one, the intercept in each model can be
interpreted as the estimated two-party preferred vote for an electorate
with mean characteristics\footnote{Mean of all variables aside from
  Judaism, Indigenous, Islam and Buddhism, where it assumes the mean of
  the log value.}. Figure \ref{fig:plotintercept} shows that the
baseline of party preference has varied over the elections, with the
biggest swing occurring in the 2007 election where the mean electorate
shifted more than five percentage points in favour of the Labor party.

\subsection{Influential
socio-demographics}\label{influential-socio-demographics}

\begin{figure}[h]

{\centering \includegraphics{draft-paper_files/figure-latex/partresplot1-1}

}

\caption{Partial residual plots by election year for a selection of predictors. Linear model with 95\% confidence bands is overlaid. Most predictors have a positive relationship: the larger the value the more likely the electorate preferences the Coalition. The relationship is relatively robust over time, with the exception of Unemployment.}\label{fig:partresplot1}
\end{figure}

\begin{figure}[h]

{\centering \includegraphics{draft-paper_files/figure-latex/partresplot2-1}

}

\caption{Partial residual plots by election year for a selection of predictors. Linear model with 95\% confidence bands is overlaid. Several predictors have a negative relationship: with larger values indicating the electorate more likely preferences Labor. Education exhibits a strong temporal change.}\label{fig:partresplot2}
\end{figure}

To investigate the socio-demographics that have a strong effect on the
two-party preferred vote, partial residual plots are used and shown in
Figures \ref{fig:partresplot1} and \ref{fig:partresplot2}. The partial
residuals are the residuals from the fitted model with the estimated
effect an individual variable added. These show the direction, size and
significance of an estimated effect --- the slope of the prediction line
matches the estimated coefficient, and the shaded region represents a
95\% confidence band. If a horizontal line can be drawn through the
confidence band, then the effect is insignificant. The estimated
intercept is also added to the partial residuals for interpretability.
Plots for each election are faceted to compare the effects over time in
Figures \ref{fig:partresplot1} and \ref{fig:partresplot2}.

It is important here to note the ecological fallacy: insights are being
drawn at the electorate level, and cannot be inferred for another
disaggregate level (e.g., individual voters).

\subsubsection{Income and unemployment}\label{income-and-unemployment}

Typically the Labor party campaigns on more progressive policies, which
often include tax reform that adversely affects higher income earners,
and more generous social assistance programs. Perhaps it is due to these
policies that higher income electorates appear more likely to support
the Liberal party, as the \texttt{Incomes} factor has a positive effect
on Liberal preference (see row 1 in Figure \ref{fig:partresplot1}). This
effect is significant in every election aside from 2004, where it is
only marginally insignificant (\(p = 0.0613\)). Unemployment however, is
not as influential. In 2001 and 2004, electorates with higher
unemployment align with Labor, but over time this shifts towards support
for the Liberal party, culminating in a significantly positive effect in
2016.

\subsubsection{Industry and type of
work}\label{industry-and-type-of-work}

Electorates with higher proportions of workers in mining, gas, water,
agriculture, waste and electricity (grouped as \texttt{Extractive}
industries) are consistently linked with higher support for the Liberal
party, with the magnitude of this effect slightly increasing over the
years (see row 3 in Figure \ref{fig:partresplot1}). This is
unsurprising, as the Liberal party has close ties with these traditional
energy industries, and typically present policies to reduce taxation on
energy production. Furthermore, electorates with more workers in
construction or manufacturing industries (\texttt{Transformative}) are
also more likely to support the Liberal party (see row 4 in Figure
\ref{fig:partresplot1}).

Similarly, the proportion of workers in managerial, administrative,
clerical and sales roles (\texttt{ManagerAdmin}) is also a significant
predictor of two-party preference vote across all six elections, with a
higher proportion of people working these jobs increasing Liberal
support. The magnitude of this effect also seems to increase over the
years.

\subsubsection{Household mobility}\label{household-mobility}

In each of the six elections, electorates with a higher proportion of
people that have recently (in the past five years) moved house
(\texttt{DiffAddress}) are more likely to support the Liberal party,
although this effect was marginally insignificant in 2013 (see row 6 in
Figure \ref{fig:partresplot1}. Having controlled for characteristics of
house ownership and rental prices (via the factors
\texttt{PropertyOwned} and \texttt{RentLoan} respectively), this effect
is somewhat surprising.

\subsubsection{Relationships}\label{relationships}

De facto relationships, but not marriages, are found to be an important
(and significant) predictor of the two-party preferred vote in all six
elections, with more de facto relationships associated with higher
support for the Labor party. The proportion of individuals who are
married however, is insignificant (not shown).

\subsubsection{Age}\label{age}

Regions comprising more older people are often believed to be more
conservative, and indeed it found that electorates with a higher median
age are more likely to support the Liberal party --- although this
effect is only significant in 2007 and 2010 (see row 2 in Figure
\ref{fig:partresplot2}).

\subsubsection{Education}\label{education}

Since 2007, electorates with higher education levels are associated with
supporting the Labor party, although this effect is only significant in
2016. Before 2007, education has an almost zero effect (see row 3 in
Figure \ref{fig:partresplot1}).

\subsubsection{Diversity}\label{diversity}

Larger migrant populations from Asia, the Middle East, South-Eastern
Europe, the United Kingdom and elsewhere, are either associated with
Labor support, or have no effect. Of these areas, only South-Eastern
European populations appear significant in each election, with the
proportion of Asian migrants also being significant in 2010. Speaking
other languages (aside from English) however, appears to have a far
stronger effect, as observed through the \texttt{OtherLanguage}
variable. Electorates with more diverse speech are associated with
higher support for the Liberal party from 2004 onwards, with this effect
being significant in 2007, 2010 and 2016. Furthermore, of the variables
relating to religion, only Judaism shows a consistent effect, with
electorates with relatively large Jewish populations more likely to vote
Liberal.

\subsubsection{A note on similar
variables}\label{a-note-on-similar-variables}

Many of the Census variables represent similar information, which is why
factors were created and some variables were removed. However, some
variables remain which are closely related. For example, an electorate's
income level (via \texttt{Incomes}) is likely to be related to electoral
unemployment and labor force participation (via \texttt{Unemployment}).
In 2001, the coefficient estimate for \texttt{Unemployment} is negative
but not significant, whilst the \texttt{Incomes} variables is
significant. If the \texttt{Incomes} variable is removed from the model
in 2001, \texttt{Unemployment} absorbs the negative effect, becoming
significant (\(p = 0.0056\)).

\subsection{A closer look at the
residuals}\label{a-closer-look-at-the-residuals}

\subsubsection{Residuals by state}\label{residuals-by-state}

It is often hypothesized that states have systematic differences that
cause their electorates to vote differently. Boxplots of residuals
grouped by state (Figure \ref{fig:resstate}) reveal that the data
reflects this -- there appears to be a state-specific effect not
captured by the models. Tasmania and the Australian Capital Territory
appear to have a bias towards Labor, whereas the Northern Territory
tends towards voting Liberal. However, there are relatively few
electorates in each of these states (five, two and two respectively), so
this apparent result may be due to incumbent effects rather than an
actual state-specific bias.

\begin{figure}[h]

{\centering \includegraphics[width=0.7\linewidth,]{draft-paper_files/figure-latex/resstate-1}

}

\caption{Boxplot of residuals by state with jittered points. States ordered by median residual. A state-specific bias not captured by the model is evident.}\label{fig:resstate}
\end{figure}

\subsubsection{Outlier electorates}\label{outlier-electorates}

Based on the distribution of the Cook's distance values, a Cook's
distance greater than \(0.1\) is considered to be influential and a
potential outlier. The electorate of Sydney (NSW) has a large Cook's
distance from 2001 to 2013, due to its diverse population (language,
birthplace and religion), high number of defacto relationships, high
income, high household mobility and small amount of workers in
extractive and transformative jobs. It has remained a strong supporter
of the Labor party and the Liberal vote is severely overpredicted by the
model, making it an outlier. Nearby in metropolitan NSW, the electorate
of Wentworth is found to be an outlier in all but the 2007 election.
Although historically Liberal, its two-party vote jumped by over 10
percentage points in 2010 without experiencing any notable changes in
its socio-demographic makeup --- implying that this may be the direct
effect of its Liberal member, Malcolm Turnbull, becoming the leader of
the Liberal party. Liberal support in Wentworth is underpredicted by the
model in each year, and more so with Turnbull as Liberal leader.

Lingiari, an electorate taking up almost all of the Northern Territory,
is an outlier in the 2001--2007 elections due to its large Indigenous
population, young age profile and low rates of property ownership.
Fowler (NSW) has a diverse population with a high proportion of
migrants, many Buddhists and Muslims, and has strong Labor support,
making it influential in 2001, 2004 and 2010. Other electorates with
large Cook's distance are Barton (NSW) and Leichhardt (QLD) in 2016, and
Canberra (ACT) in 2007.

\section{Conclusion}\label{conclusion}

This paper explores the effects of electoral socio-demographic
characteristics on the two-party preferred vote in the 2001--2016
elections, using information from the corresponding Australian federal
elections and Censuses. As a Census does not always occur in the same
year as an election, Census data for the 2004--2013 elections are
generated by employing a method of spatio-temporal imputation. This
imputes electoral socio-demographics for the electoral boundaries in
place at the time of the election --- an approach that is distinctly
different from previous work on modelling election outcomes, where
Census and election data are typically joined without addressing their
temporal differences. Before estimating a model, these socio-demographic
variables are standardized (to adjust for changing variable scales) and
many variables (representing similar information) are combined into
factors, resulting in a reduced predictor set. A spatial error model is
then estimated for each election, accounting for the inherent spatial
structure of the data.

Across the past six elections, most of the socio-demographics that drive
the electoral two-party preferred vote are found to remain steady,
whilst a few (typically weaker) effects vary over time. Industry and
type of work are particularly influential, with energy-related and
manufacturing/construction jobs, as well as administrative roles being
strongly linked with the Liberal party in all elections. Incomes have a
similarly consistent effect, with higher income areas supporting
Liberal. Higher levels of unemployment shift from weak association with
Labor to a significant Liberal effect over the years, and higher
education levels are associated with Labor from 2007 (although only
significant in 2016). It is also found that electorates with higher
household mobility support Liberal, birthplace diversity favours Labor
and more de facto relationships align with Labor preference --- although
marriages, family and household sizes have no material influence.
Furthermore, the neighbourhood (spatial) effects are found to be
positive in all elections, although only significant in 2001 and 2016,
meaning that in the 2004--2013 elections, electorates effectively voted
independently.

The findings in this paper complement the existing literature by
modelling temporal trends, which as far as the authors are aware, has
not been done previously for Australian elections using a regression
framework. It is also the first study to model any Australian election
since 2010 using Census information.

Additionally, a key contribution of this research is the wrangling of
the raw data and imputed data sets for the 2004, 2007, 2010 and 2013
elections, which have been contributed to the \texttt{eechidna}
\texttt{R} package --- providing a rich, accessible data resource for
future Australian electoral analysis.

\section{Acknowledgements}\label{acknowledgements}

This paper was produced using \texttt{RMarkdown} \autocite{rmarkdown}
and \texttt{knitr} \autocite{knitr}. All corresponding code for this
paper can be found in the github repository
\href{https://github.com/jforbes14/journal-writing}{github.com/jforbes14/journal-writing},
and the data used is available in the \texttt{eechidna} package
\autocite{eechidna}. All raw data was obtained from the Australian
Electoral Commission, the Australian Bureau of Statistics and the
Australian Government.

\section{Software}\label{software}

All election and Census datasets, along with electoral maps and more,
are available in the \texttt{eechidna} (Exploring Election and Census
Highly Informative Data Nationally for Australia) \texttt{R} package,
which can be downloaded from CRAN. The \texttt{eechidna} package makes
it easy to look at the data from the Australian Federal elections and
Censuses that occurred between 2001 and 2016. This study contributed a
large revision to the \texttt{eechidna} package, which included the
addition of election and Census data for 2001--2010, voting outcomes for
polling booths and imputed Census data for election years. For more
details on using \texttt{eechidna}, please see the articles (vignettes)
on the github page
\href{https://ropenscilabs.github.io/eechidna/}{ropenscilabs.github.io/eechidna/}.

The authors would like to sincerely thank Anthony Ebert, Heike Hofmann,
Thomas Lumley, Ben Marwick, Carson Sievert, Mingzhu Sun, Dilini
Talagala, Nicholas Tierney, Nathaniel Tomasetti, Earo Wang and Fang
Zhou, all of whom have contributed to the \texttt{eechidna} package.

\printbibliography

\end{document}